Gelation as condensation frustrated by hydrodynamics and mechanical tension

### 2017-01-13 manuscript4.tex
Describe fig:microrheology2P

\documentclass[footinbib,amsmath,amssymb,superscriptaddress,twocolumn]{revtex4}
\usepackage{graphicx,amsmath}
\usepackage{color}
\begin{document}

\title{Gelation as condensation frustrated by hydrodynamics and mechanical tension}
\author{Hideyo Tsurusawa}
\affiliation{ {Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan} }
%\thanks{These authors contributed equally to this work}
\author{Mathieu Leocmach}
\affiliation{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière, F-69622, VILLEURBANNE, France}
%\thanks{These authors contributed equally to this work}
\author{Hajime Tanaka}
%\email{tanaka@iis.u-tokyo.ac.jp}
\affiliation{ {Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan} }

@@

\begin{figure}
\includegraphics{figs/microrheology2P.pdf}
\caption{\textbf{Time-dependent microrheology} of three percolating samples.
By increasing density from top to bottom $\phi=8,\,16,\,27~\%$, $c_p=1.5,\,1.2,\, 1$ g/$\mathrm{\ell}$. 
The thick gray curve is the internal stress $\Sigma$ obtained from the measure of bond-breaking rate. 
Elastic ($G^\prime$) and viscous ($G^{\prime\prime}$) shear moduli, obtained by two-particle microrheology, are drawn respectively as red filled and black open circles. 
The dotted vertical line shows the percolation time.
}
\end{figure}

>>>>

\documentclass[footinbib,amsmath,amssymb,superscriptaddress,twocolumn]{revtex4}
\usepackage{graphicx,amsmath}
\usepackage{color}

\newcommand{\FigLossTangent}{XXX}

\begin{document}

\title{Gelation as condensation frustrated by hydrodynamics and mechanical tension}
\author{Mathieu Leocmach}
\thanks{These authors contributed equally to this work}
\affiliation{Univ Lyon, Universit\'e Claude Bernard Lyon 1, CNRS, Institut Lumi\`ere Mati\`ere, F-69622, VILLEURBANNE, France}
\author{Hideyo Tsurusawa}
\thanks{These authors contributed equally to this work}
\affiliation{ {Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan} }
\author{John Russo}
\affiliation{ {School of Mathematics, University of Bristol, Bristol BS8 1TW, United Kingdom} }
\author{Hajime Tanaka}
\email{tanaka@iis.u-tokyo.ac.jp}
\affiliation{ {Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan} }

@@

\begin{figure}
\includegraphics{figs/microrheology2P.pdf}
\caption{\textbf{Time-dependent microrheology} of three percolating samples.
By increasing density from top to bottom $\phi=8,\,16,\,27~\%$, $c_p=1.5,\,1.2,\, 1$ g/$\mathrm{\ell}$. 
The thick gray curve is the internal stress $\Sigma$ obtained from the measure of bond-breaking probability. 
Elastic ($G^\prime$) and viscous ($G^{\prime\prime}$) shear moduli at high frequency ($f=0.1\tau_\mathrm{B}^{-1}$), obtained by two-particle microrheology, are drawn respectively as red filled and black open circles. 
The dotted vertical line shows the percolation time.
}
\label{fig:microrheology2P}
\end{figure}

Following \citet{Lindstrom2012} we express the bond dissociation rate $k_D$ under a force $\mathcal{F}$ as 
\begin{equation}
k_D(\mathcal{F}) = k_D(\mathcal{F}=0) \exp\left(\frac{\mathcal{F} \delta}{k_\mathrm{B}T}\right)
\end{equation}
with $\delta$ the width of the potential, here the depletion range obtained from generalized free volume theory~\cite{Fleer2008}. $k_D(\mathcal{F}=0)$ depends only on the depth and shape of the interaction potential. From our particle-level experimental data, we can measure at each time step the probability for a bond to break. This probability decays from high values at short times to an asymptotic, finite value at long times. We checked that $k_D(t\rightarrow\infty)$ decreases with increasing polymer concentration and thus the attractive potential depth. Although, due to the small but finite value of the Debye length, $k_D(t\rightarrow\infty)$ cannot be related to the theoretical depth of the depletion potential~\cite{VanDoorn2016}.

The decay of $k_D$ takes tens of Brownian times, a time scale much longer than ion diffusion throughout the cell. Therefore the interaction potential can be considered stable and the evolution of $k_D$ translates directly in an evolution of $\mathcal{F}$. 
\begin{equation}
\mathcal{F}(t) = \frac{k_\mathrm{B}T}{\delta} \log\frac{k_D(t)}{k_D(t\rightarrow\infty)}
\end{equation}
where we assume that the final state is force-free $k_D(t\rightarrow\infty) = k_D(\mathcal{F}=0)$. Using the contact area of the depletion potential we can convert $f$ into a statistical measure of the internal stress $\Sigma (t) = 2 \mathcal{F}(t) /(\pi \sigma \delta^2)$.

To put internal stress relaxation in perspective, we need to know when the system becomes solid. The rheological definition of the gelation point is the time when the elastic and viscous parts of the shear modulus (respectively $G^\prime$ and $G^{\prime\prime}$) both scale as identical power laws of frequency $f$, in other words $G^{\prime\prime}/G^\prime$ is independent of the frequency~\cite{Chambon1987}. To measure 
the shear modulus, we use the particles themselves as passive rheological probes. Following~\citet{Crocker2000} we convert two-particle mean square displacements $\Delta r^2(t, \Delta t)$ in $G^\prime(t, f)$ and $G^{\prime\prime}(t, f)$. Supplementary figure~\FigLossTangent{} shows the evolution of $G^{\prime\prime}/G^\prime$ for three percolating samples, for three frequencies each. The convergence of the different frequencies is obvious for all samples, however only the two samples with highest volume fractions actually reach the gelation point during the experimental time. Nonetheless all three samples reach $G^\prime > G^{\prime\prime}$ at high frequencies.

In Figure~\ref{fig:microrheology2P}, to relate the timing of stress relaxation, mechanical gelation and percolation, we superimpose the time evolutions of internal stress and both parts of the shear moduli at the highest accessible frequency ($f=0.1\tau_\mathrm{B}^{-1}$). The time needed for $\Sigma(t)$ to relax completely seems to be approximatively independent of colloid volume fraction. Moreover, stress relaxation occurs independently of percolation. In dilute gels, Figure~\ref{fig:microrheology2P}a, internal stresses are relaxed when percolation takes place, whereas in dense gels, Figure~\ref{fig:microrheology2P}c, stress relaxation occurs mostly after percolation. Percolation is obviously a necessary condition for mechanical gelation, since the crossing of $G^\prime$ and $G^{\prime\prime}$ occurs always later than percolation. However there is a long delay between percolation and the rise of $G^\prime$, therefore percolation is not enough to ensure mechanical stability. In dense gel, Figure~\ref{fig:microrheology2P}c, the just percolated network rearranges under internal stresses until a mechanically stable configuration is found. For dilute gels, Figure~\ref{fig:microrheology2P}a, we observe larger fluctuations on $\Sigma(t)$ after percolation, probably due to the release of local internal stresses progressively leading to a mechanically stable configuration.

%- dense gels have much higher moduli than dilute ones
